---
title: "explore_dataVariants"
author: "FD, NOR, FB"
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.par = TRUE)
```

```{r}
today <- Sys.Date() # today's date
```

Mise à jour du `r today`.   
Source des données : <https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-pour-le-depistage-indicateurs-sur-les-variants/>


---
QUESTIONS:   
-  Ajouter données Flash ? Pour avoir des points plus anciens  
---


Légende
```{r}
library("RColorBrewer")
brw <- brewer.pal(12, "Set3")
brw[2] <- brw[12] # Darker yellow
brw[10] <- brw[11] # Lighter shade
colsAge <- c("#000000FF", brw[1:10])
names(colsAge) <- c("0", "9", "19", "29", "39", "49", "59", "69", "79", "89", "90")
pchAge <- c(16, 0:9)
names(pchAge) <- names(colsAge)
cexAge <- c(1.2, rep(1, 10))
names(cexAge) <- names(colsAge)

ages <- c("tous", "0-9", "10-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+")
```


```{r}
par(mfrow = c(1, 1))
plot(0:1, 0:1, type = "n", axes = FALSE, xlab = "", ylab = "")
legend(x = 0.5, y = 1, legend = ages, pch = pchAge, col = colsAge)
```

# Données France

## Load data 

```{r}
# Données France
URL <- "https://www.data.gouv.fr/fr/datasets/r/c43d7f3f-c9f5-436b-9b26-728f80e0fd52"
dataFile <- paste0("data/France_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.France <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

# Format date
dat.France$date1 <- as.Date(substring(dat.France$semaine, 1, 10))
dat.France$date2 <- as.Date(substring(dat.France$semaine, 12, 21))

# Rewrite time as days since beginning of the data
dat.France$time <- dat.France$date2 - min(dat.France$date2)

# Compute data on total tests
dat.France$Nb_tests <- dat.France$Nb_tests_PCR_TA_crible / (dat.France$Prc_tests_PCR_TA_crible / 100)

# Dictionnary to reformat age class
dic.age <- (0:9)*10 + 4.5 # Median of age classes
names(dic.age) <- as.character(c(0 + (0:8)*10 + 9, 90))
```

Compare the data to another source -- number of positive tests

```{r}
# Compare to another source
URL <- "https://www.data.gouv.fr/fr/datasets/r/dd0de5d9-b5a5-4503-930a-7b08dc0adc7c"
dataFile <- paste0("data/tests-France_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.tests <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

URL <- "https://www.data.gouv.fr/fr/datasets/r/c1167c4e-8c89-40f2-adb3-1954f8fedfa7"
dataFile <- paste0("data/tests7j-France_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.tests7j <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)
dat.tests7j$date1 <- as.Date(substring(dat.tests7j$semaine_glissante, 1, 10))
dat.tests7j$date2 <- as.Date(substring(dat.tests7j$semaine_glissante, 12, 21))

# P nombre de tests positifs
# T nombre de tests total

dat.tests$date <- as.Date(dat.tests$jour)
dateRange <- range(c(range(dat.France$date1), range(dat.France$date2)))

subdat.tests <- dat.tests[dat.tests$date >= dateRange[1] & dat.tests$date <= dateRange[2],]

# Function to compute a sliding window 
sliding.window <- function(v, winwdt = 7, pos = 4, na.rm = TRUE){
  # v vector to be averaged/summed
  # winwdt width of the window 
  # pos position of the focal day in the window
  # FUN function to apply
  n <- length(v)
  # Initialize output vector
  out <- 0 * v + (-1)
  out[1:(pos-1)] <- NA
  out[(n + 1 - winwdt + pos) : n] <- NA
  
  for(i in pos : (n - winwdt + pos)){
    out[i] <- mean(v[(i - pos + 1):(i + winwdt - pos)], na.rm = na.rm)
  }
  return(out[1:n])
}

subdat.tests.0 <- subdat.tests[subdat.tests$cl_age90 == 0, ]
dat.France.0 <- dat.France[dat.France$cl_age90 == 0, ]

subdat.tests.0$P.7.1 <- 7 * sliding.window(subdat.tests.0$P, pos = 1)
subdat.tests.0$P.7.4 <- 7 * sliding.window(subdat.tests.0$P, pos = 4)
subdat.tests.0$P.7.7 <- 7 * sliding.window(subdat.tests.0$P, pos = 7)
subdat.tests.0$P.8.8 <- 8 * sliding.window(subdat.tests.0$P, pos = 8, winwdt = 8)

plot(dat.France.0$date2, dat.France.0$Nb_tests, ylim = c(1*10^5, 2*10^5), pch = 16, 
     xlab = "date", ylab = "nombre de tests")
points(subdat.tests.0$date, subdat.tests.0$P.7.1, ylim = c(0, 5*10^5), col = "red")
points(subdat.tests.0$date, subdat.tests.0$P.7.4, ylim = c(0, 5*10^5), col = "green")
points(subdat.tests.0$date, subdat.tests.0$P.7.7, ylim = c(0, 5*10^5), col = "blue")
points(subdat.tests.0$date, subdat.tests.0$P.8.8, ylim = c(0, 5*10^5), col = "purple")
points(dat.tests7j$date2, dat.tests7j$P, pch = 2)
legend(x = as.Date("2021-02-18"), y = 200000, col = c("black", "red", "green", "blue", "purple", "black"), legend = c("sidep tests", "w7, c1", "w7, c4", "w7, c7", "w8, c8", "sidep 7j-fin"), pch = c(16, rep(1, 4), 2))
```

Legend notation:  
w: width of the window, c: position of the index day.  
So the sliding window is on the 7 last days.  
The difference (about 15% more positives in the variants dataset) may be due to the variant data being in terms of tests, and the other in terms of people, with duplicates removed. 

```{r}
#plot(dat.France.0$date2, dat.France.0$Nb_tests, ylim = c(1*10^5, 2*10^5), pch = 16, 
#     xlab = "date", ylab = "comparaison nombre de tests")
#points(subdat.tests.0$date, subdat.tests.0$P.7.7, ylim = c(0, 5*10^5), col = "blue")
#points(subdat.tests.0$date, 1.17*subdat.tests.0$P.7.7, ylim = c(0, 5*10^5), col = "orange")
```

Format the data further (age class data)

```{r}
# Data per age class
dat.France.ages <- dat.France[dat.France$cl_age90 != 0,]

# Add new age class code -- median of the age class
dat.France.ages$ageClass <- dic.age[as.character(dat.France.ages$cl_age90)]

# Standardize age class values
dat.France.ages$stdage <- (dat.France.ages$ageClass - mean(dat.France.ages$ageClass))/dat.France.ages$ageClass


```

## V1

### Plot
```{r, fig.height = 6}
# All ages, time
par(las = 1)
plot(dat.France$date2, dat.France$Prc_susp_501Y_V1, ylim = c(0, 100), 
     col = colsAge[as.character(dat.France$cl_age90)], 
     pch = pchAge[as.character(dat.France$cl_age90)], 
     cex = cexAge[as.character(dat.France$cl_age90)], 
     xlab = "date", ylab = "Proportion V1", axes = FALSE)
axis(1, pos = 0, at = as.Date(unique(dat.France$date2)), labels = format(unique(dat.France$date2), format = "%b %d"))
axis(2)
```

```{r, eval = FALSE}
# Plot dp/(p(1-p)) for each age class
ageClasses <- sort(unique(dat.France$cl_age90))
nAge <- length(ageClasses)

for(iage in unique(dat.France$cl_age90)){
  sub <- dat.France[dat.France$cl_age90 == iage, ]
  V1 <- sub$Prc_susp_501Y_V1/100
  t <- sub$date2
  s <- diff(V1) / (V1[-length(V1)]*(1-V1[-length(V1)]))
  plot(t[-length(V1)], s, ylim = c(-0.1, 0.1),  main = iage)
  print(c(iage, mean(s)))
}
```

### Test

Binomial model

```{r}
# Create new colums with information on number of specific PCR tests
# PCR with V1 result
dat.France.ages$V1 <- dat.France.ages$Nb_susp_501Y_V1
# All other PCRs (considering NAs are non-V1)
dat.France.ages$notV1 <- dat.France.ages$Nb_tests_PCR_TA_crible - dat.France.ages$Nb_susp_501Y_V1
# All other PCRs with a result (removing NAs)
dat.France.ages$notV1.narm <- dat.France.ages$Nb_susp_501Y_V2_3 + dat.France.ages$Nb_susp_ABS

# Check that columns correctly sum
all(dat.France.ages$Nb_susp_501Y_V2_3 + dat.France.ages$Nb_susp_ABS + dat.France.ages$Nb_susp_501Y_V1 + dat.France.ages$Nb_susp_IND - dat.France.ages$Nb_tests_PCR_TA_crible == 0)
```


#### GLM, assume that all indeterminate PCRs are non-V1

1) Age as factor, compare models without and with interaction between age and time

```{r}
# GLM
# Assuming that all IND (indetermine) are non-V1
mdl0 <- glm(cbind(V1, notV1) ~ time * factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl0)

# Without interaction
mdl1 <- glm(cbind(V1, notV1) ~ time + factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl1)

## Likelihood ratio test
anova(mdl1, mdl0, test="Chisq")
```

2) Age as quantitative

```{r}
# Test date effect
mdl2 <- glm(cbind(V1, notV1) ~ time * ageClass, data = dat.France.ages, family = "binomial")
summary(mdl2)

## Likelihood ratio test
anova(mdl2, mdl0, test="Chisq")
```

#### GLM, ignore indeterminate PCRs

1) Age as factor, compare models without and with interaction between age and time

```{r}
# GLM
# Assuming that all IND (indetermine) are non-V1
mdl0.narm <- glm(cbind(V1, notV1.narm) ~ time * factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl0.narm)

# Without interaction
mdl1.narm <- glm(cbind(V1, notV1.narm) ~ time + factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl1.narm)

## Likelihood ratio test
anova(mdl1.narm, mdl0.narm, test="Chisq")
```

2) Age as quantitative

```{r}
# Test date effect
mdl2.narm <- glm(cbind(V1, notV1.narm) ~ time * ageClass, data = dat.France.ages, family = "binomial")
summary(mdl2.narm)

## Likelihood ratio test
anova(mdl2.narm, mdl0.narm, test="Chisq")
```


## V2/V3

### Plot
```{r, fig.height = 6}
# All ages, time
par(las = 1)
plot(dat.France$date2, dat.France$Prc_susp_501Y_V2_3, ylim = c(0, 100), 
     col = colsAge[as.character(dat.France$cl_age90)], 
     pch = pchAge[as.character(dat.France$cl_age90)], 
     cex = cexAge[as.character(dat.France$cl_age90)], 
     xlab = "date", ylab = "Proportion V2/V3", axes = FALSE)
axis(1, pos = 0, at = as.Date(unique(dat.France$date2)), labels = format(unique(dat.France$date2), format = "%b %d"))
axis(2)
```

### Test 

Binomial 

```{r}
# Create new colums with information on number of specific PCR tests
# PCR with V2/V3 result
dat.France.ages$V23 <- dat.France.ages$Nb_susp_501Y_V2_3
# All other PCRs (considering NAs are non-V23)
dat.France.ages$notV23 <- dat.France.ages$Nb_tests_PCR_TA_crible - dat.France.ages$Nb_susp_501Y_V2_3
# All other PCRs with a result (removing NAs)
dat.France.ages$notV23.narm <- dat.France.ages$Nb_susp_501Y_V1 + dat.France.ages$Nb_susp_ABS
```

#### GLM, assume that all indeterminate PCRs are non-V1

1) Age as factor, compare models without and with interaction between age and time

```{r}
# GLM
# Assuming that all IND (indetermine) are non-V23
mdl0 <- glm(cbind(V23, notV23) ~ time * factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl0)

# Without interaction
mdl1 <- glm(cbind(V23, notV23) ~ time + factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl1)

## Likelihood ratio test
anova(mdl1, mdl0, test="Chisq")
```

2) Age as quantitative

```{r}
# Test date effect
mdl2 <- glm(cbind(V23, notV23) ~ time * ageClass, data = dat.France.ages, family = "binomial")
summary(mdl2)

## Likelihood ratio test
anova(mdl2, mdl0, test="Chisq")
```

#### GLM, ignore indeterminate PCRs

1) Age as factor, compare models without and with interaction between age and time

```{r}
# GLM
# Assuming that all IND (indetermine) are non-V23
mdl0.narm <- glm(cbind(V23, notV23.narm) ~ time * factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl0.narm)

# Without interaction
mdl1.narm <- glm(cbind(V23, notV23.narm) ~ time + factor(ageClass), data = dat.France.ages, family = "binomial")
summary(mdl1.narm)

## Likelihood ratio test
anova(mdl1.narm, mdl0.narm, test="Chisq")
```

2) Age as quantitative

```{r}
# Test date effect
mdl2.narm <- glm(cbind(V23, notV23.narm) ~ time * ageClass, data = dat.France.ages, family = "binomial")
summary(mdl2.narm)

## Likelihood ratio test
anova(mdl2.narm, mdl0.narm, test="Chisq")
```

## Multinomial 

Frequencies

```{r}
par(mfrow = c(1, 1))
x <- dat.France.0$date2
c1 <- dat.France.0$Prc_susp_501Y_V1
c23 <- c1 + dat.France.0$Prc_susp_501Y_V2_3
cABS <- c23 + dat.France.0$Prc_susp_ABS
cIND <- cABS + dat.France.0$Prc_susp_IND

colV <- c(brewer.pal(3, "Set2"), gray(0.5))

plot(x, c1, ylim = c(0, 100), type = "l", 
     xlab = "date", ylab = "Proportion variants", axes = FALSE)
axis(1, pos = 0, labels = format(x, "%b-%d"), at = x)
axis(2, pos = x[1])
polygon(c(x, rev(x)), c(rep(0, length(x)), rev(c1)), col = colV[1], border = NULL)
polygon(c(x, rev(x)), c(c1, rev(c23)), col = colV[2], border = NULL)
polygon(c(x, rev(x)), c(c23, rev(cABS)), col = colV[3], border = NULL)
polygon(c(x, rev(x)), c(cABS, rev(cIND)), col = colV[4], border = NULL)

nend <- length(x)-1
text(x = rep(x[nend], 4), y = c(c1[nend]/2, c1[nend] + (c23[nend] - c1[nend])/2, c23[nend] + (cABS[nend] - c23[nend])/2, cABS[nend] + (cIND[nend] - cABS[nend])/2), labels = c("V1", "V2/V3", "WT", "indet."), col = "white", adj = 1)

lines(dat.France.0$date2, dat.France.0$Prc_susp_501Y_V1 + dat.France.0$Prc_susp_501Y_V2_3)
lines(dat.France.0$date2, dat.France.0$Prc_susp_501Y_V1 + dat.France.0$Prc_susp_501Y_V2_3 )
lines(dat.France.0$date2, dat.France.0$Prc_susp_501Y_V1 + dat.France.0$Prc_susp_501Y_V2_3)
```


```{r}
mat <- as.matrix(data.frame(WT = dat.France.ages$Nb_susp_ABS, 
                            V1 = dat.France.ages$Nb_susp_501Y_V1, 
                            V2 = dat.France.ages$Nb_susp_501Y_V2_3, 
                            INDET = dat.France.ages$Nb_susp_IND))

library(nnet)

## Null model
m0 <- multinom(mat ~ 1)
summary(m0)

## Compute by hand to check result
log(colSums(mat)[2]/colSums(mat)[1])
log(colSums(mat)[3]/colSums(mat)[1])
log(colSums(mat)[4]/colSums(mat)[1])

## Time effect
m1 <- multinom(mat ~ time, data=dat.France.ages)
summary(m1)

## Likelihood ratio test
anova(m0, m1, test="Chisq")

## Age effect
m2 <- multinom(mat ~ time + stdage, data=dat.France.ages)
summary(m2)

## Likelihood ratio test
anova(m1, m2, test="Chisq")

## Age effect
m3 <- multinom(mat ~ time * stdage, data=dat.France.ages)
summary(m3)

## Likelihood ratio test
anova(m2, m3, test="Chisq")
```

Pb in the data is the change in the age structure of positive individuals; to really test the effet, we would need information on the negative tests as well, and the age distribution in these negative tests. 

```{r}
tapply(dat.France.ages$Nb_tests_PCR_TA_crible, list(dat.France.ages$cl_age90, dat.France.ages$time), sum)
```


# Données régions

## Load data 

```{r}
URL <- "https://www.data.gouv.fr/fr/datasets/r/73e8851a-d851-43f8-89e4-6178b35b7127"
dataFile <- paste0("data/Regions_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.regions <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

# Format date
dat.regions$date1 <- as.Date(substring(dat.regions$semaine, 1, 10))
dat.regions$date2 <- as.Date(substring(dat.regions$semaine, 12, 21))

# Rewrite time as days since beginning of the data
dat.regions$time <- dat.regions$date2 - min(dat.regions$date2)

# Compute data on total tests
dat.regions$Nb_tests <- dat.regions$Nb_tests_PCR_TA_crible / (dat.regions$Prc_tests_PCR_TA_crible / 100)
```

```{r}
# Codes regions
URL <- "https://www.data.gouv.fr/en/datasets/r/34fc7b52-ef11-4ab0-bc16-e1aae5c942e7"
dataFile <- "data/coderegions.csv"
download.file(URL, dataFile)
codesRegions <- read.csv(dataFile, sep = ",", stringsAsFactors = FALSE)

# Turn into dictionary
regs <- codesRegions$nom_region
names(regs) <- as.character(codesRegions$code_region)
```

```{r}
# Add region name
dat.regions$reg_name <- regs[as.character(dat.regions$reg)]
# dat.regions[floor(runif(10)*1000), c("reg", "reg_name")] # check a few names

# What are the other regions??
# aggregate(dat.regions$reg, by = list(dat.regions$reg), FUN = length)
```

Format data further

```{r}
dat.regions.ages <- dat.regions[dat.regions$cl_age90 != 0,]

# Add new age class code -- median of the age class
dat.regions.ages$ageClass <- dic.age[as.character(dat.regions.ages$cl_age90)]

# Standardize age class values
dat.regions.ages$stdage <- (dat.regions.ages$ageClass - mean(dat.regions.ages$ageClass))/dat.regions.ages$ageClass

```

## V1

### Plot

```{r fig.height = 12}
tmp <- unique(dat.regions$reg) # Region codes
tmp <- tmp[tmp>10 & tmp <= 93] # Choose only metropolitan regions
par(mfrow = c(4, 3))
for(region in tmp){
  subdat <- dat.regions[dat.regions$reg == region, ]
  plot(subdat$date2, subdat$Prc_susp_501Y_V1, ylim = c(0, 100), main = regs[as.character(region)], col = colsAge[as.character(subdat$cl_age90)], pch = pchAge[as.character(subdat$cl_age90)], 
       xlab = "date", ylab = "Proportion V1"
       )
}
```

### Test

```{r}
# Create new colums with information on number of specific PCR tests
# PCR with V1 result
dat.regions.ages$V1 <- dat.regions.ages$Nb_susp_501Y_V1
# All other PCRs (considering NAs are non-V1)
dat.regions.ages$notV1 <- dat.regions.ages$Nb_tests_PCR_TA_crible - dat.regions.ages$Nb_susp_501Y_V1
# All other PCRs with a result (removing NAs)
dat.regions.ages$notV1.narm <- dat.regions.ages$Nb_susp_501Y_V2_3 + dat.regions.ages$Nb_susp_ABS

# Check that columns currently sum
all(dat.regions.ages$Nb_susp_501Y_V2_3 + dat.regions.ages$Nb_susp_ABS + dat.regions.ages$Nb_susp_501Y_V1 + dat.regions.ages$Nb_susp_IND - dat.regions.ages$Nb_tests_PCR_TA_crible == 0)

dat.regions.ages$reg_name.fac <- as.factor(dat.regions.ages$reg_name)
```

#### GLM, assume that all indeterminate PCRs are non-V1

1) Age as factor, compare models without and with interaction between age and time

```{r}
# GLM
# Assuming that all IND (indetermine) are non-V1
mdl0 <- glm(cbind(V1, notV1) ~ time * factor(ageClass) * factor(reg_name), data = dat.regions.ages, family = "binomial")
summary(mdl0)

# Without interaction
mdl1 <- glm(cbind(V1, notV1) ~ time*factor(reg_name) + factor(ageClass), data = dat.regions.ages, family = "binomial")
summary(mdl1)

# Without interaction
mdl1bis <- glm(cbind(V1, notV1) ~ time*factor(ageClass) + factor(reg_name), data = dat.regions.ages, family = "binomial")
summary(mdl1bis)

# Without interaction at all
mdl2 <- glm(cbind(V1, notV1) ~ time + factor(ageClass) + factor(reg_name), data = dat.regions.ages, family = "binomial")
summary(mdl2)

## Likelihood ratio tests
anova(mdl1, mdl0, test="Chisq")
anova(mdl1bis, mdl0, test="Chisq")
anova(mdl2, mdl0, test="Chisq")
```

2) Age as quantitative

```{r}
mdl3 <- glm(cbind(V1, notV1) ~ time * ageClass * factor(reg_name), data = dat.regions.ages, family = "binomial")
summary(mdl3)


## Likelihood ratio test
anova(mdl3, mdl0, test="Chisq")
```

```{r}
# GLM
# Assuming that all IND (indetermine) are non-V1
mdl <- glm(cbind(V1, notV1) ~ date2 * as.factor(ageClass) + reg_name.fac, data = dat.regions.ages, family = "binomial")
summary(mdl)

# Ignoring IND
mdl.narm <- glm(cbind(V1, notV1.narm) ~ date2 + cl_age90 + reg_name.fac, data = dat.regions.ages, family = "binomial")
summary(mdl.narm)
```

## V2 

```{r fig.height = 12}
par(mfrow = c(4, 3))
for(region in tmp){
  subdat <- dat.regions[dat.regions$reg == region, ]
  plot(subdat$date2, subdat$Prc_susp_501Y_V2_3, ylim = c(0, 100), main = regs[as.character(region)], col = colsAge[as.character(subdat$cl_age90)], pch = pchAge[as.character(subdat$cl_age90)], 
       xlab = "date", ylab = "Proportion V2/V3"
       )
}
```

```{r}
# Create new colums with information on number of specific PCR tests
# PCR with V2/3 result
dat.regions.ages$V23 <- dat.regions.ages$Nb_susp_501Y_V2_3
# All other PCRs (considering NAs are non-V1)
dat.regions.ages$notV23 <- dat.regions.ages$Nb_tests_PCR_TA_crible - dat.regions.ages$Nb_susp_501Y_V2_3
# All other PCRs with a result (removing NAs)
dat.regions.ages$notV23.narm <- dat.regions.ages$Nb_susp_501Y_V1 + dat.regions.ages$Nb_susp_ABS

# GLM
# Assuming that all IND (indetermine) are non-V23
mdl <- glm(cbind(V23, notV23) ~ date2 + ageClass + reg_name.fac, data = dat.regions.ages, family = "binomial")
summary(mdl)

# Ignoring IND
mdl.narm <- glm(cbind(V23, notV23.narm) ~ date2 + ageClass + reg_name.fac, data = dat.regions.ages, family = "binomial")
summary(mdl.narm)

```

# Départements

```{r}
URL <- "https://www.data.gouv.fr/fr/datasets/r/16f4fd03-797f-4616-bca9-78ff212d06e8"
dataFile <- paste0("data/Dep_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.deps <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

# Format date
dat.deps$date1 <- as.Date(substring(dat.deps$semaine, 1, 10))
dat.deps$date2 <- as.Date(substring(dat.deps$semaine, 12, 21))

```

```{r}
# Add name
deps <- read.csv("data/departement2020.csv", stringsAsFactors = FALSE)
# Turn into dictionnary
dps <- deps$libelle
names(dps) <- as.character(deps$dep)
```

```{r}
dat.deps$departement <- dps[as.character(dat.deps$dep)]
```

## V1

```{r figsDepartements1, fig.height = 100, fig.width = 7}
par(mfrow = c(35, 3))
for(idep in sort(unique(dat.deps$dep))){
  tmp <- dat.deps[dat.deps$dep == idep, ]
  plot(tmp$date2, tmp$Prc_susp_501Y_V1, ylim = c(0, 100), col = colsAge[as.character(tmp$cl_age90)], pch = pchAge[as.character(tmp$cl_age90)], 
       xlab = "date", ylab = "Proportion V1", 
       main = unique(tmp$departement))
  
#  legend(x = min(dat.deps$date2), y = 100, legend = ages, pch = pchAge, col = colsAge)
}
```

## V2/V3

```{r figsDepartements, fig.height = 100, fig.width = 7}
par(mfrow = c(35, 3))
for(idep in sort(unique(dat.deps$dep))){
  tmp <- dat.deps[dat.deps$dep == idep, ]
  plot(tmp$date2, tmp$Prc_susp_501Y_V2_3, ylim = c(0, 100), col = colsAge[as.character(tmp$cl_age90)], pch = pchAge[as.character(tmp$cl_age90)], 
       xlab = "date", ylab = "Proportion V2/V3", 
       main = unique(tmp$departement))
  
#  legend(x = min(dat.deps$date2), y = 100, legend = ages, pch = pchAge, col = colsAge)
}
```

